\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-30 metatheory-condensed.tex
add linkage rules in selected-typing.tex

As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are represented by de Bruijn indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Typing judgments for substitutions have the form 
%\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (e.g., rule \ruleref{tm/sub}).
%
The two main forms of substitutions are weakening and extension:
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.

@@

In contrast, $\wsig$ should be hidden from the typing context of any
term that does not invoke $\Wsup{\wsig}{\cdot}{\cdot}$ or $\Wrec{\wsig}{\cdot}{\cdot}$, so that
the term can be reused---without being rechecked---for a different
W-type signature~$\wsig'$ that extends~$\wsig$ with additional
constructors.
Moreover, the typing of the term should be made polymorphic to
the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
so that the term can be reused---without being rechecked---when those fields
are overridden to support the extended signature~$\wsig'$.

>>>>

As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are represented by de Bruijn indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Typing judgments for substitutions have the form 
%\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (e.g., rule \ruleref{tm/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.

@@

In contrast, $\wsig$ should be hidden from the typing context of any
term that does not invoke $\Wsup{\wsig}{\cdot}{\cdot}$ or $\Wrec{\wsig}{\cdot}{\cdot}$, so that
the term can be reused---without being rechecked---for a different
W-type signature~$\wsig'$ that extends~$\wsig$ with additional
constructors.
Moreover, the typing of the term should be made parametric to
the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
so that the term can be reused---without being rechecked---when those fields
are overridden to support the extended signature~$\wsig'$.
%
Such abstraction required by family polymorphism is supported via linkages.

\noindentparagraph{Family polymorphism via linkages.}

Family polymorphism requires late binding.
In \TT, families are expressed through linkages, and late binding of
field references is achieved by requiring that typing be polymorphic to
the definition of that field.
